Geometric variational problems

### 2020-11-02 NSF2020/Description/specificaim1_ver2.tex
...

\section{Proposed Research}
\label{sec:proposed-work}
The goal of the proposed research is to develop fast,
high-order-accurate, parallel numerical algorithms for large-scale
simulations of the collective hydrodynamics of  amphiphilic particles in a viscous solvent.
%
We have demonstrated that 
our potential theory approach can efficiently simulate self-assembly of 
amphiphilic particles into two-dimensional micelles, bilayer membranes, and vesicles \cite{Fu19}.
%
While these results show great potentials in simulating the collective hydrodynamics of amphiphilic particles and
reproducing mechanical properties of their bilayer assembly, 
several outstanding issues need to be addressed for such approach to be efficiently applied to three-dimensional 
collective hydrodynamics of amphiphilic particles.

@@

 We conjecture that 
once assembled, the bilayer shapes possess some or all of the mechanical properties of bilayers from continuum theory.

>>>>

\section{Proposed Research}
\label{sec:proposed-work}
The proposed research develops fast,
high-order-accurate, parallel numerical algorithms for large-scale
simulations of the collective hydrodynamics of amphiphilic particles in a viscous solvent.
%
We have demonstrated that our potential approach efficiently simulates
self-assembly of amphiphilic particles into two-dimensional micelles, bilayer membranes, and vesicles \cite{Fu19}, and
recreates the tank-treading phenomenon in external shear flows \cite{Fu20}.
%
While the results show great promise in the field of collective body hydrodynamics,
several outstanding issues need to be addressed. These include a thorough 
analysis of potentially new mechanical properties possessed by HAP bilayers, and 
efficiently simulating three-dimensional collective hydrodynamics of amphiphilic particles.
Finally, we must mathematically characterize the variational behavior of HAP under certain homogenized limits.

@@

Hamm and Kozlov (HK) pioneered the modern theory of membrane continuum mechanics in the early 2000s \cite{HaKo2000}.
The HK framework assumes a three-dimensional, internal structure for lipid monolayers and 
then uses the theory of elasticity to derive an effective, two-dimensional elastic energy.
The internal structure consists of straight fibers that represent elongated hydrocarbon chains, and these fibers
tilt and stretch with respect to the dividing surface. The elastic theory assumes an energy density that is
quadratic in the Green-Lagrange strain tensor. Due to the internal structure ansatz, the elastic energy decomposes into
four, fundamental and independent deformations: splay, saddle splay, twist and tilt.
The elastic moduli for each deformation comes from a transversal moment of the internal stress.
These moduli define the mechanical properties of lipid monolayers and bilayers.

(Importance, take from Akimov PRE and Zimmerberg Soft Matter)

Recently, there has been a revival in interest in the HK theory. The HK framework assumes a quadratic
elasticity energy density and this has caused researchers in molecular dynamics to question the validity of
the continuum theory \cite{PRL}. Zimmerberg and Akimov \cite{}, however, showed that for simplified deformations.
that molecular dynamics and continuum theory agree even at small length scales.

The elastic energy of monolayer is a function of the material coordinate and its derivatives. 
That is because the changes in free energy, in elasticity theory, come from changes in spacing between mass points in a material.
There are a number of mathematical models of membrane elasticity in applied mathematics. Our HAP formulation, however,
is the first to explicitly consider intermolecular repulsion, and hydrophobic interaction with the solvent 
to calculate the free energy of a lipids in a monolayer.

What is the mathematical definition of a lipid monolayer and a deformation? 
The reference state will lipids lying on a flat surface with their tails pointing in the 
vertical direction. For concreteness, take the reference region to be the triangular prism 
\begin{equation}
\mathcal{R}_0 = \{(X_1, X_2, X_3) \mid 0 \leq X_1, 0 \leq X_2, X_1 + X_2 \leq a, 0 \leq X_3 \leq \delta \}
\end{equation}
The parameter $\delta$ is the monolayer height and $a$ is the edge length for the triangular base $\mathcal{T}$. 
The size and shape of the base is arbitrary.

The deformed state maps the flat base $\mathcal{T}$ to a curved surface $\mathcal{S} = x_0(\mathcal{T})$. 
The vertical line segments in $\mathcal{R}_0$ represent
hydrocarbon chains and map to line segments  in the deformed state.
A director field $\mathbf{n}(\mathbf{x})$ for $\mathbf{x} \in \mathcal{S}$ specifies the direction. 
Based on these considerations, the deformation takes the form  
\begin{equation}
x(X_1, X_2, X_3) = x_0(X_1, X_2) + \zeta(x_0(X_1, X_2), X_3) n(x_0(X_1, X_2))
\end{equation}    
To satisfy incompressibility, the chains stretch and compress parallel to their axis, but do not form kinks.
For a point $P \in \mathcal{S},$ the function $\zeta(P, X_3)$ parametrizes the changes in change length and 
$n(P)$ is the chain axis.  
To simplify matters, we perform the calculations at one point on the surface. 
Thus, assume, without loss of generality, that $x_0(0,0) = (0,0,0) = O$.

The field of membrane continuum mechanics still lacks concensus as to the form of the elastic energy when
it involves complicated deformations. Recently, Deserno included a tilt gradient term that was proportedly incorrectly neglected from the HK theory.
Later Akimov and Deserno independently realized that there was inconsitency regarding the transversal invariance
of tilt in Desrno's earlier work. The latest revision of HK theory by Akimov contains more than ten independent
contributions to the elastic energy. We have found that the explosion in the number of elastic terms
is a result of expanding the energy density in an incorrect reference frame, and that using a natural
reference frame leads to a simplified energy resembling the Frank theory from the study of liquid crystals.

Our investigations will leave open the possibility that the free energy of monolayers and bilayers is non-local. 
We have shown that hydrophic attraction is a non-additive force, meaning the force of attraction between two
particles is dependent on the position a third or more particles. Conceptually, this would arise in continuum theory if two monolayer surfaces
had different elastic energy density at that point despite having locally identical curvatures and directors.

Importance of the theory

Mathematical setup and past contributions
(old material)

Revision

We conjecture that the monolayer shapes 
possess some or all of the mechanical properties of bilayers from continuum theory.
